Pairwise-complete correlation considered dangerous

### 2015-05-26 missing.Rmd
missing

The functions `cov(x)` computes the symmetric covariance matrix `V` with entries defined by the pairwise covariance of columns of `x`,
```{r,eval=FALSE}
V[i,j] = cov(x[,i], x[,j])
```
where `i=j=c(1,2,3)` in this example. The function `cor(x)` similarly computes the symmetric correlation matrix with entries defined by pairwise correlation of the columns of `x`. For example:
```{r}
cov(x)
cor(x)
```
Due to missing values in the third column of `x` we know that we don't know the covariance between `x[,3]` and anything else. Thanks to an arguably questionable^[The `cov2cor` always puts ones along the diagonal; that choice is valid only if all unknowns may assume bounded numeric values.] choice in R's `cov2cor` function, we see that `cor(x[,3],x[,3])=1`, but we don't know the correlation between `x[,3]` and the other columns.

>>>>

The functions `cov(x)` computes the symmetric covariance matrix `V` with entries defined by the pairwise covariance of columns of `x`,
```{r,eval=FALSE}
V[i,j] = crossprod(x[,i] - mean(x[,i]), x[,j] - mean(x[,j])) / nrow(x)
```
where `i=j=c(1,2,3)` in this example. The function `cor(x)` similarly computes the symmetric correlation matrix with entries defined by pairwise correlation of the columns of `x`. For example:
```{r}
cov(x)
cor(x)
```
Due to missing values in the third column of `x` we know that we don't know the covariance between `x[,3]` and anything else. Thanks to an arguably questionable^[The `cov2cor` always puts ones along the diagonal; that choice is valid only if all unknowns may assume bounded numeric values.] choice in R's `cov2cor` function, we see that `cor(x[,3],x[,3])=1`, but we don't know the correlation between `x[,3]` and the other columns.
